An analysis of Lenin's work: Revolution, war, and fruta

### 2013-12-21 paper/distance.tex
Add another work to documents being analyzed.

where $f_w$ is the frequency of appearance of $w$ in $T$. This value captures, in some way, whether a word
has a distribution in text that is not exactly uniform; an order of the word in the text that brings some sense to it.

Having these value, we can pull the most representative words out of a text by selecting those of greater information value. Now our question arises: how distant is a word from a text? We use here the notion of distance between a point and a set defined as:

@@

\begin{itemize}
  \item What is to be done(1901)
  \item Materialism and Empiriocriticism (1908)
  \item Imperialism, the highest stage of capitalism (1916)
  \item State and Revolution (1917)
\end{itemize}

@@

@@

\section{Results}
\subsection{Resultados sobre obras}
* Tabla de top words de Estado y Revolución
\begin{center}
  \begin{tabular}{ | l | l | l |}
    \hline
    \multicolumn{3}{|c|}{Top words for `State and Revolution'} \\ \hline
    \# & Word & Information Value \\ \hline

    1 & kautsky & 0.10 \\ \hline
    2 &  marx & 0.09 \\ \hline
    3 &  revolution & 0.09 \\ \hline
    4 &  works & 0.06 \\ \hline 
    5 & engels & 0.06 \\ \hline
    6 &  society & 0.06 \\ \hline
    7 & power & 0.05 \\ \hline
    8 &  party & 0.04 \\ \hline
    9 &  social & 0.04 \\ \hline
    10 &  experience & 0.03 \\ \hline 
    11 & bernstein & 0.03 \\ \hline
    12 &  question & 0.03 \\ \hline 
    13 & equality & 0.03 \\ \hline
    14 &  phase & 0.03 \\ \hline
    15 & 1917 & 0.03 \\ \hline
    16 & commune & 0.03 \\ \hline
    17 & one & 0.03 \\ \hline
    18 & republic & 0.03 \\ \hline
    19 & class & 0.02 \\ \hline
    20 & law & 0.02 \\ \hline
  \end{tabular}
\end{center}

\begin{center}
  \begin{tabular}{ | r |l |}
    \hline
    \multicolumn{2}{|c|}{$S_{dw}($`State and Revolution'$, w)$} \\ \hline
      $w$ & $S_{dw}(d, w)$ \\ \hline
      revolution &  0.177  \\ \hline
      power &  0.169  \\ \hline
      state &  0.154  \\ \hline
      commune &  0.133  \\ \hline
      republic &  0.123  \\ \hline
      politics &  0.118  \\ \hline
      music &  0.114  \\ \hline
      war &  0.110  \\ \hline
      philosophy &  0.094  \\ \hline
      idealism &  0.084  \\ \hline
      chicken &  0.092  \\ \hline
      computer &  0.082  \\ \hline
      mosquito &  0.044  \\ \hline
      lion &  0.059  \\ \hline
      alcohol &  0.067  \\ \hline
      toilette &  0.069  \\ \hline
  \end{tabular}
\end{center}

@@

\begin{thebibliography}{10}
\bibitem{DARWIN}
M.A. Montemurro and D. H. Zanette, {\em The statistics of meaning: Darwin, Gibbon and Moby Dick}, Significance, Dec. 2009, 165-169.
\bibitem{ENTROPIC}
M.A. Montemurro and D. H. Zanette, {\em Entropic analysis of the role of words in literary texts}, Adv. Complex Systems 5, September 27th 2001
\end{thebibliography}
\end{article}

>>>>

where $f_w$ is the frequency of appearance of $w$ in $T$. This value captures, in some way, whether a word
has a distribution in text that is not exactly uniform; an order of the word in the text that brings some sense to it. Ordering the words by their information value results in a list of the text's most representative words.

In order to measure the distance/similarity between two words, we use Wordnet (see \cite{WORDNET}), a lexical database in English. Using a subset of all the possible ``senses'' of both words , we test several similarity notions between them.

Having a similarity notion between words, and the most valuable words from a text, our question arises: how distant is a word from a text? We use here the notion of distance between a point and a set defined as:

@@

\begin{itemize}
  \item What is to be done(1901)
  \item The Agrarian Programme of Social-Democracy in the First Russian Revolution, 1905-1907 (1907)
  \item Materialism and Empiriocriticism (1908)
  \item Imperialism, the highest stage of capitalism (1916)
  \item State and Revolution (1917)
\end{itemize}

@@

\subsection{Top words for given works}

\begin{center}
  \begin{tabular}{ | l | l | }
    \hline
    \multicolumn{2}{|c|}{State and Revolution} \\ \hline
    \# & Word  \\ \hline
1 & kautsky \\ \hline
2 & marx \\ \hline
3 & revolution \\ \hline
4 & works \\ \hline
5 & engels \\ \hline
6 & society \\ \hline
7 & power \\ \hline
8 & party \\ \hline
9 & social \\ \hline
10 & experience \\ \hline
11 & bernstein \\ \hline
12 & question \\ \hline
13 & equality \\ \hline
14 & phase \\ \hline
15 & 1917 \\ \hline
16 & commune \\ \hline
17 & one \\ \hline
18 & republic \\ \hline
19 & class \\ \hline
20 & law \\ \hline
  \end{tabular}
\end{center}

\begin{center}
  \begin{tabular}{ | l | l | }
    \hline
    \multicolumn{2}{|c|}{Imperialism, the Highest ...} \\ \hline
    \# & Word  \\ \hline
1 & enterprises \\ \hline
2 & bank \\ \hline
3 & imperialism \\ \hline
4 & world \\ \hline
5 & million \\ \hline
6 & colonies \\ \hline
7 & kautsky \\ \hline
8 & colonial \\ \hline
9 & production \\ \hline
10 & cent \\ \hline
11 & economic \\ \hline
12 & company \\ \hline
13 & capitalist \\ \hline
14 & monopoly \\ \hline
15 & big \\ \hline
16 & countries \\ \hline
17 & policy \\ \hline
18 & powers \\ \hline
19 & 000 \\ \hline
20 & banks \\ \hline
  \end{tabular}
\end{center}

\begin{center}
  \begin{tabular}{ | l | l | }
    \hline
\multicolumn{2}{|c|}{Materialism and Empirio-criticism} \\ \hline
\# & Word  \\ \hline
1 & physics \\ \hline
2 & social \\ \hline
3 & avenarius \\ \hline
4 & space \\ \hline
5 & time \\ \hline
6 & dietzgen \\ \hline
7 & berkeley \\ \hline
8 & truth \\ \hline
9 & rey \\ \hline
10 & objective \\ \hline
11 & sensation \\ \hline
12 & experience \\ \hline
13 & pp \\ \hline
14 & engels \\ \hline
15 & motion \\ \hline
16 & marx \\ \hline
17 & book \\ \hline
18 & works \\ \hline
19 & vol \\ \hline
20 & general \\ \hline
  \end{tabular}
\end{center}

\begin{center}
  \begin{tabular}{ | l | l | }
    \hline
\multicolumn{2}{|c|}{What Is To Be Done?} \\ \hline
\# & Word  \\ \hline
1 & local \\ \hline
2 & organisation \\ \hline
3 & abroad \\ \hline
4 & union \\ \hline
5 & criticism \\ \hline
6 & economic \\ \hline
7 & political \\ \hline
8 & revolutionaries \\ \hline
9 & freedom \\ \hline
10 & newspapers \\ \hline
11 & workers \\ \hline
12 & nadezhdin \\ \hline
13 & consciousness \\ \hline
14 & newspaper \\ \hline
15 & bourgeois \\ \hline
16 & masses \\ \hline
17 & martynov \\ \hline
18 & would \\ \hline
19 & movement \\ \hline
20 & german \\ \hline
  \end{tabular}
\end{center}

\begin{center}
  \begin{tabular}{ | l | l | }
    \hline
\multicolumn{2}{|c|}{The Agrarian Programme of ...} \\ \hline
\# & Word  \\ \hline
1 & rent \\ \hline
2 & dessiatins \\ \hline
3 & 000 \\ \hline
4 & land \\ \hline
5 & maslov \\ \hline
6 & peasants \\ \hline
7 & capital \\ \hline
8 & revolution \\ \hline
9 & marx \\ \hline
10 & people \\ \hline
11 & nationalisation \\ \hline
12 & plekhanov \\ \hline
13 & restoration \\ \hline
14 & right \\ \hline
15 & latifundia \\ \hline
16 & private \\ \hline
17 & farming \\ \hline
18 & free \\ \hline
19 & central \\ \hline
20 & lands \\ \hline
  \end{tabular}
\end{center}

@@

@@

\begin{thebibliography}{10}
\bibitem{DARWIN}
M.A. Montemurro and D. H. Zanette, {\em The statistics of meaning: Darwin, Gibbon and Moby Dick}, Significance, Dec. 2009, 165-169.
\bibitem{ENTROPIC}
M.A. Montemurro and D. H. Zanette, {\em Entropic analysis of the role of words in literary texts}, Adv. Complex Systems 5, September 27th 2001
\bibitem{WORDNET}
George A. Miller, {\em WordNet: A Lexical Database for English }, COMMUNICATIONS OF THE ACM November 1995/Vol. 38, No. 11

\end{thebibliography}
\end{article}
